Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-26 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
cleaning eft appendix

The theory admits as a solution, the garden-variety Schwarzschild BH of GR.
This is not the case when rotation is included and the Kerr metric is not a
solution of the theory~\cite{Jackiw:2003pm}.
%
These rotating BHs support a scalar field which fall off as
$r^{-2}$ asymptotically, to which we can associate a scalar dipole
charge~\cite{Yunes:2009hc,Konno:2009kg} and
%
the leading-order modification to the GW phase enters at 2PN~\cite{Yagi:2011xp}.
%
Deviations from GR at this PN order are constrained with present GW observations~\cite{LIGOScientific:2021sio}
only at the level of $\sim 50\%$ (see the constraint on the 2PN parameter $\varphi_4$ in Fig.~6 of
Ref.~\cite{LIGOScientific:2021sio}). So far, analyses that used only the inspiral portion of
the BBH GW signals were not able to set meaningful bounds on such deviation at 2PN order~\cite{Nair:2019iur,Perkins:2021mhb}.
%
Nonetheless, the theory has been constrained in Ref.~\cite{Silva:2020acr}, which found
%
$\ell_{\rm dCS} \leqslant 8.5$~km,
%
by folding data from the X-ray observations of the isolated NS
PSR~J0030+0451~\cite{Lommen:2000yt,NANOGrav:2017wvv} by
NICER~\cite{Riley:2019yda,Miller:2019cac} and from the GW observation of the binary
NS GW170817~\cite{TheLIGOScientific:2017qsa,LIGOScientific:2018cki}, using
equation-of-state independent relations between the NS moment of inertia and tidal deformability~\cite{Yagi:2013bca,Yagi:2013awa,Gupta:2017vsl}.

The QNMs of the Schwarzschild BH in dCS gravity were studied in
Refs.~\cite{Yunes:2007ss,Cardoso:2009pk,Molina:2010fb}, which found that scalar
perturbations couple to gravitational perturbations of axial parity, in contrast
with EdGB gravity.
%
This leads to a breakdown of isospectrality. The QNM spectra of slowly-rotating BHs in dCS gravity was
studied in Ref.~\cite{Srivastava:2021imr,Wagle:2021tam}.
%
They were also extracted from NR simulations of BH head-on collisions in Ref.~\cite{Okounkova:2019dfo}.

@@

More specifically, we follow the notation of
Refs.~\cite{Cano:2020cao,Cano:2021myl} and consider up to dimension-eight
operators ($n=4$),
%
\begin{subequations}
\label{eq:action_mods}
\begin{align}
    L^{(6)} &= \lame R_{\mu\nu}{}^{\rho\sigma} R_{\rho\sigma}{}^{\gamma\delta} R_{\gamma \delta}{}^{\mu\nu}
    + \lamo R_{\mu\nu}{}^{\rho\sigma} R_{\rho\sigma}{}^{\gamma\delta} \tilde{R}_{\gamma \delta}{}^{\mu\nu}\,,
    \label{eq:s_6}
    \\
    L^{(8)} &= \varepsilon_{1} \mathcal{C}^2
    + \varepsilon_{2} \tilde{\mathcal{C}}^{2}
    + \varepsilon_{3} \mathcal{C} \tilde{\mathcal{C}}\,,
\label{eq:s_8}
\end{align}
\end{subequations}
%
where ${\cal C} = R_{\mu\nu\rho\sigma} R^{\mu\nu\rho\sigma}$,
$\tilde{\cal C} = R_{\mu\nu\gamma\delta} \epsilon^{\mu\nu}{}_{\rho\sigma} R^{\rho\sigma\gamma\delta}$,
and both $\lambda_{\rm o, e}$ and $\varepsilon_{i}$ (with $i=1, 2, 3$) are dimensionless parameters.
%
Due to the large number of free parameters in this theory, we focus on a subset of the parameter space.
%
In particular, we consider dimension-six and dimension-eight operators separately.
%
In addition, in the dimension-six case we further assume that $\lame = \lamo = 1$,
leaving us with $\ell_{\rm cEFT}$ as our single free parameter.
%
Similarly, in the dimension-eight case, we set $\varepsilon_{1} = 1$ and $\varepsilon_{2} = \varepsilon_{3} = 0$,
as done in Ref.~\cite{Sennett:2019bpc}.
%
This leaves us with $\ell_{\rm qEFT}$ as our single free parameter.

@@

\begin{table*}[t]
% \begin{tabular}{c | c c c c c c c c c c}
% \hline
% \hline
% Theory      & $\omega^{(0)}_{220}$ & $\tau^{(0)}_{220}$ & $\omega^{(1)}_{220}$ & $\tau^{(1)}_{220}$ & $p$ & $\delta \omega^{(0)}_{220}$ & $\delta \tau^{(0)}_{220}$ & $\delta \omega^{(1)}_{220}$ & $\delta \tau^{(1)}_{220}$ & Ref.  \\
% \hline
% GR          & 0.3737               & 11.2407            &  0.1258              & 0.2522             & --  & --                          & --                        & --                          & --                        & \cite{Maselli:2019mjd} \\
% \hline
% EdGB        & --                   & --                 &  --                  & --                 & 4   & 0.0107                      & 0.0044                    & $-$0.2480                   & $-$1.101                  & \cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} \\
% dCS         & --                   & --                 &  --                  & --                 & 4   & 3.1964                      & 6.3619                    & 41.199                      & 794.66                    & \cite{Wagle:2021tam}  \\
% cubic EFT   & --                   & --                 &  --                  & --                 & 4   & $-$0.5813                   & 2.6469                    & $-$3.8620                   & 265.12                    & \cite{Cano:2021myl} \\
% quartic EFT & --                   & --                 &  --                  & --                 & 6   & $-$0.2114                   & $-$0.6070                 & $-$1.5263                   & 171.35                    & \cite{Cano:2021myl}  \\
% \hline
% \hline
% \end{tabular}
%
\begin{tabular}{c | c }
\hline
\hline
                        & GR~\cite{Maselli:2019mjd} \\
\hline
$\omega^{(0)}_{220}$    & 0.3737                    \\
$\tau^{(0)}_{220}$      & 11.2407                   \\
$\omega^{(1)}_{220}$    & 0.1258                    \\
$\tau^{(1)}_{220}$      & 0.2522                    \\
\hline
\hline
\end{tabular}
\quad
\begin{tabular}{c | c  c  c  c}
\hline
\hline
                              & EdGB ($p=4$)~\cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} & dCS ($p=4$)~\cite{Wagle:2021tam} & cubic EFT ($p=4$)~\cite{Cano:2021myl} & quartic EFT ($p=6$)~\cite{Cano:2021myl} \\
\hline
$\delta\omega^{(0)}_{220}$    & 0.0107                                               & 3.1964                   & $-$0.5813                     & $-$0.2114                       \\
$\delta\tau^{(0)}_{220}$      & 0.0044                                               & 6.3619                   & $-$0.2114                     & $-$0.6070                       \\
$\delta\omega^{(1)}_{220}$    & $-$0.2480                                            & 41.199                   & $-$3.8620                     & $-$1.5263                       \\
$\delta\tau^{(1)}_{220}$      & $-$1.1010                                            & 794.66                   & 265.12                        & 171.35                          \\
\hline
\hline
\end{tabular}
\caption{Summary of theory-specific QNM calculations.
%
We summarize each theory we have considered together with: the exponent $p$ at
which their QNM-modification enters, the corresponding modifications to the
oscillation frequency $\delta \omega^{(n)}_{220}$ and decay time $\delta \tau^{(n)}_{220}$, and the
references from which we used the results from.
%
We also included for comparison the GR coefficients, up to $j=1$, obtained in Ref.~\cite{Maselli:2019mjd}.
%
The remaining GR coefficients for $1< j \leqslant 4$, for which their non-GR counterpart cannot be determined as of yet for
the theories under consideration, can be found in Table~I of Ref.~\cite{Maselli:2019mjd}.
}
\label{tab:ref_theories_qnms}
\end{table*}

How can we determine the beyond-GR corrections? In GR, comparison between the numerically determined
Kerr QNMs against the fitting formula~\eqref{eq:qnm_nongr_iso} fixes the GR expansion coefficients
$\omega^{(j)}$ and $\tau^{(j)}$.
%
We can proceed in a similar way with QNMs calculated in the context of a non-GR theory.
%
In particular, in the literature, we can already find fitting formulas relating
the QNMs to the BHs mass, spin and length-scale $\ell_{\rm th}$, the latter being
specific to each theory, up to $j=1$ in the spin expansion (see Table~\ref{tab:ref_theories_qnms}).
%
The idea is then to compare these formulas against Eq.~\eqref{eq:qnm_nongr_iso}
to fix $p$, $\delta\omega^{(j)}$, and $\delta\tau^{(j)}$.
%
Because QNMs of rotating BHs in modified gravity theories are not known to
all spin values, we can expect that the $j=1$ coefficients to be susceptible to
changes as calculations beyond-leading order in spin are accomplished in the future.
%
That is not the case for the $j=0$ coefficients and the situation is the same as in GR,
in which the $j=0$ coefficients are simply the QNMs of the Schwarzschild BH.

@@

There are two conditions that we must verify before we can confidently claim to
have placed a constraint on $\ell_{\rm th}$. First, as we have explained in Sec.~\ref{sec:review_theories}, all theories that we
consider must be interpreted as an EFT, meaning that it should be
considered valid only below an energy scale, or equivalently, a length-scale.
%
As a cut-off length-scale for the validity of the EFT we use,
%
\begin{equation}
\Lambda_{\rm EFT} (\varepsilon, \mathfrak{m}) = \varepsilon \, \frac{G \mathfrak{m}}{c^2} \,,
\label{eq:def_cutoff}
\end{equation}
%
where $\varepsilon$ is a dimensionless number and $\mathfrak{m}$ is the median
value of one of the mass scales involved in the problem.
%
We note that $\Lambda_{\rm EFT}$ has dimensions of length and hence can be compared to each theory's fundamental length-scale $\ell_{\rm th}$.
%
Here we explore the range $\varepsilon \in [0, 1]$, but following
Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr} we quote our final
results using $\varepsilon = 1/2$, \hs{but we stress that there is no
fundamental justification for using this cut off.}
% %
% In Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}, $\varepsilon = 1/2$ was used,
% but here we explore $\varepsilon \in [0, 1]$ for generality.
% %

@@

Let us now consider the quartic EFT of GR, as our final example.
%
In Fig.~\ref{fig:qEFT_exec_sum} we show the posteriors on $\ell_{\rm qEFT}$ for GW150914 (top panel),
GW200129 (middle panel) for $N_{\rm max} = 0$, which are qualitatively similar to the cubic EFT of GR.
%
We find that while the Parspec bound is satisfied, the EFT bound is only marginally so,
%
As shown in Fig.~\ref{fig:qEFT_cdf}, the 90\% credible level is reached for $\varepsilon \approx 0.58$
(in the case of GW15094) and for $\varepsilon \approx 0.64$ (in the case of GW200129).
%
Having in mind that the hard cut-off $\varepsilon = 1/2$ is not fundamental, but to keep consistency
across our analysis, our final result
%
\begin{equation}
    \ell_{\rm qEFT} \leqslant 51.3~\textrm{km},
    %\quad \textrm{at 90\% credible level},
\end{equation}
%
at 90\% credible level should be taken lightly.
%
However, we \emph{can} claim the validity of the bound above, but at a lower,
68\% credible level.

% We find that both the EFT~\eqref{eq:eft_bound} and ParSpec~\eqref{eq:parspec_bound} are satisfied,
% hence allowing us to infer bounds on $\ell_{\rm qEFT}$.
% %
% As Fig.~\ref{fig:qEFT_cdf} shows the CDFs for both events reach 90\% for $\gm = M_{\rm f}$ before $\varepsilon = 1/2$.
% %
% Consequently, from the joint posterior of $\ell_{\rm qEFT}$ we quote our final result,
% %
% \begin{equation}
%     \ell_{\rm qEFT} \leqslant 51.3~\textrm{km} \quad \textrm{at 90\% credible level}\,.
% \end{equation}
% %
% % \abcomm{Please update the paragraph below.}
In this theory, we have considered only $N_{\rm max} = 0$.
%
We find that the addition of spin corrections (while maintaining the same prior
ranges on $\ell_{\rm qEFT}$ as used in the $N_{\rm max} = 0$ study) can result
in waveforms that can have a ringdown segment larger (sometimes seconds long)
than the inspiral-plunge segment in the detectors frequency band, making the
parameter estimation challenging.
% \abcomm{The reference to 80 km comes out of the blue. Doesn't it? Why did you refer to it?}~\hscomm{That was the upper prior limit for the GW150914 run. I modified the text to explain this.}
% for which the ringdown portions of the larger than the inspiral-plunge part. \abcomm{The last sentence
% has some problems in English. I do not udnerstand it.}
%
To overcome this issue we have lowered the value of $\ell^{\rm max}_{\rm qEFT}$, but by doing so we have obtained posteriors
which were flat, just as our prior, and were thus uninformative. Hence, we do not quote any results for $N_{\rm max} = 1$.
%
Table~\ref{tab:summary_qeft} summarizes our findings for the quartic EFT of GR.
%
\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/qeft_posteriors_combined_corrected.pdf}
\caption{(Color online).~Posterior distribution function for the coupling constant $\ell_{\rm qEFT}$ for
GW150914 (top panel), GW200129 (middle panel) and combined events (bottom panel).
}
\label{fig:qEFT_exec_sum}
\end{figure}

@@

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
$N_{\rm max}$ & Event & EFT    & ParSpec & Constraint ($\gm = M_{\rm f}$) \\
              &       & bound? & bound?  &                            \\
\hline
  & GW150914 & Yes & Yes  & $\ell_{\rm qEFT} \leqslant 51.7$~km \\
0 & GW200129 & Yes & Yes  & $\ell_{\rm qEFT} \leqslant 54.8$~km \\
  & Combined &     &      & \cellcolor{black!10}$\ell_{\rm qEFT} \leqslant 51.3$~km \\
% \hline
%   & GW150914 &  &     &    \\
% 1 & GW200129 &  &     &    \\
%   & Combined &  &     &    \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results the quartic EFT of GR for GW150914, GW200129, and
combined events using $\gm = M_{\rm f}$, $\varepsilon = 1/2$, and $N_{\rm max} = 0$. The quoted
result correspond \hs{to 90\% credible values \emph{if} we allow for a more flexible cut-off $\varepsilon \lesssim 0.65$.
However, the result is robust under the hard cut-off of $\varepsilon = 1/2$ at 65\% credible level.}
}
\label{tab:summary_qeft}
\end{table}

@@

The QNMs of slowly-rotating BHs in both cubic and quartic EFT of GR where calculated in Ref.~\cite{Cano:2021myl}.
%
For the cubic EFT, we use their Eq.~(67), in the particular case of $\lame = \lamo = 1$.
%
%
We then linearize the resulting expression in $\chi$ and consider $m=2$ the harmonic.
%
\hscomm{Instead calling them ``$+$'' and ``$-$'' following the notation used in~\cite{Cano:2021myl},
I call them axial/polar to follow our conventions.}
%
As an outcome, we find that the fundamental axial-parity QNM is the least damped one, and it is
the one we use.
%
Direct comparison with Eqs.~\eqref{eq:qnm_nongr_iso} results in $p_{\rm cEFT} = 4$,
%
\begin{align}
    \delta \omega_{\rm cEFT}^{(0)} = - 0.5813, \quad \delta \tau_{\rm cEFT}^{(0)} = 2.6469,
    \nonumber \\
    \delta \omega_{\rm cEFT}^{(1)} = -3.8620, \quad \delta \tau_{\rm cEFT}^{(1)} = 265.12.
\end{align}

We proceed in the same away for the quartic EFT. Here we use Eq.~(68) (with
$\epsilon_{1} = 1$ and $\epsilon_{2} = 0$) and Eq.~(70) of
Ref.~\cite{Cano:2021myl},
%
In this case we find that both modes (``$-$'' and ``$+$'' in the notation of Ref.~\cite{Cano:2020cao}) make the
fundamental QNM damp fast, and we choose the ``$-$'' mode which decreases less the damping time. \abcomm{Not clear.
Decreases less with respect to what? Do you mean it has a smaller damping time?}
%
This time we then find that $p_{\rm qEFT} = 6$,
%
\begin{align}
    \delta \omega_{\rm qEFT}^{(0)} = -0.2114, \quad \delta \tau_{\rm qEFT}^{(0)} = -0.6070,
    \nonumber \\
    \delta \omega_{\rm qEFT}^{(1)} = -1.5263, \quad \delta \tau_{\rm qEFT}^{(1)} = 171.35.
\end{align}
%
As with the previous theories, the larger values of some of these parameters are compensated
by the assumptions of weak coupling and small spin used to calculate the QNM frequencies.
\
\section{The estimation of  intrinsic binary parameters in General Relativity and modified theories of gravity}
\label{app:big_contour}

>>>>

The theory admits as a solution the garden-variety Schwarzschild BH of GR.
This is not the case when rotation is included and the Kerr metric is not a
solution of the theory~\cite{Jackiw:2003pm}.
%
These rotating BHs support a scalar field which fall off as
$r^{-2}$ asymptotically, to which we can associate a scalar dipole
charge~\cite{Yunes:2009hc,Konno:2009kg} and
%
the leading-order modification to the GW phase enters at 2PN~\cite{Yagi:2011xp}.
%
Deviations from GR at this PN order are constrained with present GW observations~\cite{LIGOScientific:2021sio}
only at the level of $\sim 50\%$ (see the constraint on the 2PN parameter $\varphi_4$ in Fig.~6 of
Ref.~\cite{LIGOScientific:2021sio}). So far, analyses that used only the inspiral portion of
the BBH GW signals were not able to set meaningful bounds on such deviation at 2PN order~\cite{Nair:2019iur,Perkins:2021mhb}.
%
Nonetheless, the theory has been constrained in Ref.~\cite{Silva:2020acr}, which found
%
$\ell_{\rm dCS} \leqslant 8.5$~km,
%
by folding data from the X-ray observations of the pulsar
PSR~J0030+0451~\cite{Lommen:2000yt,NANOGrav:2017wvv} by
NICER~\cite{Riley:2019yda,Miller:2019cac} and from the GW observation of the binary
NS GW170817~\cite{TheLIGOScientific:2017qsa,LIGOScientific:2018cki}, using
equation-of-state independent relations between NS moment of inertia and tidal deformability~\cite{Yagi:2013bca,Yagi:2013awa,Gupta:2017vsl}.

The QNMs of the Schwarzschild BH in dCS gravity were studied in
Refs.~\cite{Yunes:2007ss,Cardoso:2009pk,Molina:2010fb}, which found that scalar
perturbations couple to gravitational perturbations of axial parity, in contrast
with EdGB gravity, resulting in a breakdown of isospectrality.
%
The QNM spectra of slowly-rotating BHs in dCS gravity was
studied in Refs.~\cite{Srivastava:2021imr,Wagle:2021tam}.
%
They were also extracted from NR simulations of BH head-on collisions in Ref.~\cite{Okounkova:2019dfo}.

@@

More specifically, we follow the notation of
Refs.~\cite{Cano:2020cao,Cano:2021myl} and consider up to dimension-eight
operators ($n=4$),
%
\begin{subequations}
\label{eq:action_mods}
\begin{align}
    L^{(6)} &= \lame R_{\mu\nu}{}^{\rho\sigma} R_{\rho\sigma}{}^{\gamma\delta} R_{\gamma \delta}{}^{\mu\nu}
    + \lamo R_{\mu\nu}{}^{\rho\sigma} R_{\rho\sigma}{}^{\gamma\delta} \tilde{R}_{\gamma \delta}{}^{\mu\nu},
    \label{eq:s_6}
    \\
    L^{(8)} &= \varepsilon_{1} \mathcal{C}^2
    + \varepsilon_{2} \tilde{\mathcal{C}}^{2}
    + \varepsilon_{3} \mathcal{C} \tilde{\mathcal{C}},
\label{eq:s_8}
\end{align}
\end{subequations}
%
where ${\cal C} = R_{\mu\nu\rho\sigma} R^{\mu\nu\rho\sigma}$,
$\tilde{\cal C} = R_{\mu\nu\gamma\delta} \epsilon^{\mu\nu}{}_{\rho\sigma} R^{\rho\sigma\gamma\delta}$,
and both $\lambda_{\rm o, e}$ and $\varepsilon_{i}$ (with $i=1, 2, 3$) are dimensionless parameters.
%
Due to the large number of free parameters in this theory, we focus on a subset of the parameter space.
%
In particular, we consider dimension-six and dimension-eight operators separately.
%
In addition, in the dimension-six case we further assume that $\lame = \lamo = 1$,
leaving us with $\ell_{\rm cEFT}$ as our single free parameter.
%
Similarly, in the dimension-eight case, we set $\varepsilon_{1} = 1$ and $\varepsilon_{2} = \varepsilon_{3} = 0$,
as done in Ref.~\cite{Sennett:2019bpc}.
%
This leaves us with $\ell_{\rm qEFT}$ as our single free parameter.

@@

\begin{table*}[t]
% \begin{tabular}{c | c c c c c c c c c c}
% \hline
% \hline
% Theory      & $\omega^{(0)}_{220}$ & $\tau^{(0)}_{220}$ & $\omega^{(1)}_{220}$ & $\tau^{(1)}_{220}$ & $p$ & $\delta \omega^{(0)}_{220}$ & $\delta \tau^{(0)}_{220}$ & $\delta \omega^{(1)}_{220}$ & $\delta \tau^{(1)}_{220}$ & Ref.  \\
% \hline
% GR          & 0.3737               & 11.2407            &  0.1258              & 0.2522             & --  & --                          & --                        & --                          & --                        & \cite{Maselli:2019mjd} \\
% \hline
% EdGB        & --                   & --                 &  --                  & --                 & 4   & 0.0107                      & 0.0044                    & $-$0.2480                   & $-$1.101                  & \cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} \\
% dCS         & --                   & --                 &  --                  & --                 & 4   & 3.1964                      & 6.3619                    & 41.199                      & 794.66                    & \cite{Wagle:2021tam}  \\
% cubic EFT   & --                   & --                 &  --                  & --                 & 4   & $-$0.5813                   & 2.6469                    & $-$3.8620                   & 265.12                    & \cite{Cano:2021myl} \\
% quartic EFT & --                   & --                 &  --                  & --                 & 6   & $-$0.2114                   & $-$0.6070                 & $-$1.5263                   & 171.35                    & \cite{Cano:2021myl}  \\
% \hline
% \hline
% \end{tabular}
%
\begin{tabular}{c | c }
\hline
\hline
                        & GR~\cite{Maselli:2019mjd} \\
\hline
$\omega^{(0)}$    & 0.3737                    \\
$\tau^{(0)}$      & 11.2407                   \\
$\omega^{(1)}$    & 0.1258                    \\
$\tau^{(1)}$      & 0.2522                    \\
\hline
\hline
\end{tabular}
\quad
\begin{tabular}{c | c  c  c  c}
\hline
\hline
                              & EdGB ($p=4$)~\cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} & dCS ($p=4$)~\cite{Wagle:2021tam} & cubic EFT ($p=4$)~\cite{Cano:2021myl} & quartic EFT ($p=6$)~\cite{Cano:2021myl} \\
\hline
$\delta\omega^{(0)}$    & 0.0107                                               & 3.1964                   & $-$0.5813                     & $-$0.2114                       \\
$\delta\tau^{(0)}$      & 0.0044                                               & 6.3619                   & $-$0.2114                     & $-$0.6070                       \\
$\delta\omega^{(1)}$    & $-$0.2480                                            & 41.199                   & $-$3.8620                     & $-$1.5263                       \\
$\delta\tau^{(1)}$      & $-$1.1010                                            & 794.66                   & 265.12                        & 171.35                          \\
\hline
\hline
\end{tabular}
\caption{Summary of theory-specific QNM calculations.
%
We summarize each theory we have considered together with: the exponent $p$ at
which their QNM-modification enters, the corresponding modifications to the
oscillation frequency $\delta \omega^{(n)}_{220}$ and decay time $\delta \tau^{(n)}_{220}$, and the
references from which we used the results from.
%
We also included for comparison the GR coefficients, up to $j=1$, obtained in Ref.~\cite{Maselli:2019mjd}.
%
The remaining GR coefficients for $1< j \leqslant 4$, for which their non-GR counterpart cannot be determined as of yet for
the theories under consideration, can be found in Table~I of Ref.~\cite{Maselli:2019mjd}.
%
\hscomm{I added the values of $p$ to the column headers to keep both columns with the same number of rows. I also
removed the subscript 220 because we already said we consider only this mode.}
}
\label{tab:ref_theories_qnms}
\end{table*}

How can we determine the beyond-GR corrections? In GR, comparison between the numerically determined
Kerr QNMs against the fitting formula~\eqref{eq:qnm_nongr_iso} fixes the GR expansion coefficients
$\omega^{(j)}$ and $\tau^{(j)}$.
%
We can proceed in a similar way with QNMs calculated in the context of a non-GR theory.
%
In particular, in the literature, we can already find fitting formulas relating
the QNMs to the BHs mass, spin and length-scale $\ell_{\rm th}$, the latter being
specific to each theory, up to $j=1$ in the spin expansion (see Table~\ref{tab:ref_theories_qnms}).
%
The idea is then to compare these formulas against Eq.~\eqref{eq:qnm_nongr_iso}
to fix $p$, $\delta\omega^{(j)}$, and $\delta\tau^{(j)}$.
%
Because QNMs of rotating BHs in modified gravity theories are not known to
all spin values, we can expect that the $j=1$ coefficients to change as
calculations beyond-leading order in spin are accomplished in the future.
%
That is not the case for the $j=0$ coefficients and the situation is the same as in GR,
in which the $j=0$ coefficients are simply the QNMs of the Schwarzschild BH.

@@

There are two conditions that we must verify before we can confidently claim to
have placed a constraint on $\ell_{\rm th}$. First, as we have explained in Sec.~\ref{sec:review_theories}, all theories that we
consider must be interpreted as an EFT, meaning that it should be
considered valid only below an energy scale, or equivalently, a length-scale.
%
As a cut-off length-scale for the validity of the EFT we use,
%
\begin{equation}
\Lambda_{\rm EFT} (\varepsilon, \mathfrak{m}) = \varepsilon \, \frac{G \mathfrak{m}}{c^2} \,,
\label{eq:def_cutoff}
\end{equation}
%
where $\varepsilon$ is a dimensionless number and $\mathfrak{m}$ is the median
value of one of the mass scales involved in the problem.
%
We note that $\Lambda_{\rm EFT}$ has dimensions of length and hence can be compared to each theory's fundamental length-scale $\ell_{\rm th}$.
%
Here we explore the range $\varepsilon \in [0, 1]$, but following
Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr} we quote our final
results using $\varepsilon = 1/2$, \hs{but we stress that there is no
fundamental justification for this choice.}
% %
% In Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}, $\varepsilon = 1/2$ was used,
% but here we explore $\varepsilon \in [0, 1]$ for generality.
% %

@@

Let us now consider the quartic EFT of GR, as our final example.
%
In Fig.~\ref{fig:qEFT_exec_sum} we show the posteriors on $\ell_{\rm qEFT}$ for GW150914 (top panel),
GW200129 (middle panel) for $N_{\rm max} = 0$, which are qualitatively similar to the cubic EFT of GR.
%
We find that while the Parspec bound is satisfied, the EFT bound is only marginally so,
%
As shown in Fig.~\ref{fig:qEFT_cdf}, the 90\% credible level is reached for $\varepsilon \approx 0.58$
(in the case of GW15094) and for $\varepsilon \approx 0.64$ (in the case of GW200129).
%
\hs{Having in mind that the cut off $\varepsilon = 1/2$ is not fundamental, but to keep consistency
across our analysis, our final result
%
\begin{equation}
    \ell_{\rm qEFT} \leqslant 51.3~\textrm{km},
    %\quad \textrm{at 90\% credible level},
\end{equation}
%
at 90\% credible level should be taken lightly.}
%
However, we \emph{can} claim the validity of the bound above, but at a lower,
68\% credible level.

% We find that both the EFT~\eqref{eq:eft_bound} and ParSpec~\eqref{eq:parspec_bound} are satisfied,
% hence allowing us to infer bounds on $\ell_{\rm qEFT}$.
% %
% As Fig.~\ref{fig:qEFT_cdf} shows the CDFs for both events reach 90\% for $\gm = M_{\rm f}$ before $\varepsilon = 1/2$.
% %
% Consequently, from the joint posterior of $\ell_{\rm qEFT}$ we quote our final result,
% %
% \begin{equation}
%     \ell_{\rm qEFT} \leqslant 51.3~\textrm{km} \quad \textrm{at 90\% credible level}\,.
% \end{equation}
% %
% % \abcomm{Please update the paragraph below.}
In this theory, we have considered only $N_{\rm max} = 0$.
%
We find that the addition of spin corrections (while maintaining the same prior
ranges on $\ell_{\rm qEFT}$ as used in the $N_{\rm max} = 0$ study) can result
in waveforms that can have a ringdown segment larger (sometimes seconds long)
than the inspiral-plunge segment in the detectors' frequency band, making the
parameter estimation challenging.
% \abcomm{The reference to 80 km comes out of the blue. Doesn't it? Why did you refer to it?}~\hscomm{That was the upper prior limit for the GW150914 run. I modified the text to explain this.}
% for which the ringdown portions of the larger than the inspiral-plunge part. \abcomm{The last sentence
% has some problems in English. I do not udnerstand it.}
%
To overcome this issue we have lowered the value of $\ell^{\rm max}_{\rm qEFT}$, but by doing so we have obtained posteriors
which were flat, just as our prior, and were thus uninformative. Hence, we do not quote any results for $N_{\rm max} = 1$.
%
Table~\ref{tab:summary_qeft} summarizes our findings for the quartic EFT of GR.
%
\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/qeft_posteriors_combined_corrected.pdf}
\caption{(Color online).~Posterior distribution function for the coupling constant $\ell_{\rm qEFT}$ for
GW150914 (top panel), GW200129 (middle panel) and combined events (bottom panel).
}
\label{fig:qEFT_exec_sum}
\end{figure}

@@

\begin{table}[t]
\begin{tabular}{l l c c c}
\hline
\hline
$N_{\rm max}$ & Event & EFT    & ParSpec & Constraint ($\gm = M_{\rm f}$) \\
              &       & bound? & bound?  &                            \\
\hline
  & GW150914 & Yes & Yes  & $\ell_{\rm qEFT} \leqslant 51.7$~km \\
0 & GW200129 & Yes & Yes  & $\ell_{\rm qEFT} \leqslant 54.8$~km \\
  & Combined &     &      & \cellcolor{black!10}$\ell_{\rm qEFT} \leqslant 51.3$~km \\
% \hline
%   & GW150914 &  &     &    \\
% 1 & GW200129 &  &     &    \\
%   & Combined &  &     &    \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results the quartic EFT of GR for GW150914, GW200129, and
combined events using $\gm = M_{\rm f}$, $\varepsilon = 1/2$, and $N_{\rm max} = 0$. The quoted
result correspond \hs{to 90\% credible values \emph{if} we allow for a more flexible cut-off $\varepsilon \lesssim 0.65$.
However, the result is robust under the hard cut-off of $\varepsilon = 1/2$ at 65\% credible level.}
}
\label{tab:summary_qeft}
\end{table}

@@

The QNMs of slowly-rotating BHs in both cubic and quartic EFT of GR where calculated in Ref.~\cite{Cano:2021myl}.
%
For the cubic EFT, we use their Eq.~(67), in the particular case of $\lame = \lamo = 1$.
%
%
We then linearize the resulting expression in $\chi$ and consider $m=2$ the harmonic.
%
\hscomm{Instead calling them ``$+$'' and ``$-$'' following the notation used in~\cite{Cano:2021myl},
I call them axial/polar to follow our conventions.}
%
As an outcome, we find that the fundamental axial-parity QNM is the least damped one, and it is
the one we use.
%
Direct comparison with Eqs.~\eqref{eq:qnm_nongr_iso} results in $p_{\rm cEFT} = 4$,
%
\begin{align}
    \delta \omega_{\rm cEFT}^{(0)} = - 0.5813, \quad \delta \tau_{\rm cEFT}^{(0)} = 2.6469,
    \nonumber \\
    \delta \omega_{\rm cEFT}^{(1)} = -3.8620, \quad \delta \tau_{\rm cEFT}^{(1)} = 265.12,
\end{align}
%
for this theory.

We proceed in the same away for the quartic EFT. Here we use Eq.~(68) (with
$\epsilon_{1} = 1$ and $\epsilon_{2} = 0$) and Eq.~(70) of
Ref.~\cite{Cano:2021myl},
%
In this case we find that both axial and polar modes reduce the damping time
of the fundamental QNM mode relative to GR.
%
Hence, we choose the axial-parity mode for this which reduction is the smallest.
% \abcomm{Not clear.
% Decreases less with respect to what? Do you mean it has a smaller damping time?}
%
This time we then find that $p_{\rm qEFT} = 6$,
%
\begin{align}
    \delta \omega_{\rm qEFT}^{(0)} = -0.2114, \quad \delta \tau_{\rm qEFT}^{(0)} = -0.6070,
    \nonumber \\
    \delta \omega_{\rm qEFT}^{(1)} = -1.5263, \quad \delta \tau_{\rm qEFT}^{(1)} = 171.35,
\end{align}
%
for this theory.
%
As in the case of the previous theories, the large values of some of these
coefficients are compensated by the assumptions of weak coupling and small spin
used to calculate the QNM frequencies.

\section{The estimation of  intrinsic binary parameters in General Relativity and modified theories of gravity}
\label{app:big_contour}

% HS: I agree with all the changes below
We show in Fig.~\ref{fig:corner_plot_all} a corner plot for all the intrinsic
binary parameters from our parameter-estimation study of GW150914, using the
\pSEOB{} waveform models for GR and dCS.
%
We see that the medians of the posterior distributions are not affected substantially by the inclusion of the non-GR parameters.
%
This is particularly important for the $\Mf$ and $\chi_{\rm f}$ parameters, since
the ParSpec framework assumes that the non-GR theory induces small deviations from GR.
%
\hs{Indeed, since \pSEOB{} introduces only minimal modifications to the
inspiral and because the remnant BH parameters are estimated according to GR
predictions, the fact that only small biases are introduced on $\Mf$ and $\chi_{\rm f}$
is, to some extend, expected.}
%
We obtain qualitative similar results for GW200129 and the other modified gravity theories considered in our work.
%
We also remark that the fact that the posterior on $\chi_{\rm f}$ has most support around $\approx 0.7$ justifies our use
of the fitting coefficients in the ParSpec formulas of Ref.~\cite{Maselli:2019mjd}.
%
%
% %
% \hscomm{Should we say that this happens almost by ``construction''? That would be because \pSEOB{} waveform changes very little the inspiral
% and the final masses and spins are predicted according to a GR formula.}
